Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-25 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

\subsection{Optical Torque}
It has long been known that electromagnetic fields can transfer linear and
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this project we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

@@

Birefringence is a material property often seen in crystalline materials, 
if the crystal lattice has different refractive indices for varying 
polarisations of light. For circularly polarised light this inhomogeneity 
results in a high degree of SAM being transferred to the target object \cite{Parkin2009, Arita2016}, this has been exploited to rotate microspheres 
as fast as 1000 Hz while suspended in a bulk medium \cite{Arita2016} as well 
as a means of measuring the local temperature and shear response of said 
medium \cite{Millen2014, RodriguezSevilla2018}. Calculating the optical 
torque applied to a birefringent material is given via:
\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi 
		\\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\theta$ is the angle between the particle's long axis and the 
polarisation vector of the local EM field, and $\phi$ is the phase shift 
in the EM field. The first term represents the 'orientational' torque which 
is due to the target particle being aligned with the EM amplitude, when 
aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. For sphere's this is always the case as they lack a 
distinctive major axis. The second term is due purely to the polarisation 
of the optical trap, for circularly polarised light $\phi=\pi/4$ thus 
maximising the torque transferred to the target particle. Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
birefringence, but there are other mechanisms that result in optical torque.

>>>>

\subsection{Optical Torque and rotation}
\label{sec:opt_torque}
It has long been known that electromagnetic fields can transfer linear and
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this project we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

@@

Birefringence is a material property often seen in crystalline materials, 
if the crystal lattice has different refractive indices for varying 
polarisations of light. Because of this the material has two different 
refractive indices depending on the polarisation of light. For circularly 
polarised light this inhomogeneity results in a high degree of SAM being 
transferred to the target object \cite{Parkin2009, Arita2016}. The greater 
the difference between the two refractive indces the greater the angular
momentum transfer.

The ability to transfer angular momentum has been exploited to rotate 
microspheres as fast as 1000 Hz while suspended in a bulk medium 
\cite{Arita2016} as well as a means of measuring the local temperature 
and shear response of said medium \cite{Millen2014, RodriguezSevilla2018}. 
Calculating the optical torque applied to a birefringent material is given via:
\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi 
		\\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\Delta n$ is the difference between the two refractive indices, 
$\theta$ is the angle between the particle's long axis and the polarisation 
vector of the local EM field, and $\phi$ is the phase shift in the EM 
field. The first term represents the 'orientational' torque which is 
due to the target particle being aligned with the EM amplitude, when 
aligned $\theta=0$ meaning the entire term is negligible for particle's 
with a stable orientation. For sphere's this is always the case as they 
lack a distinctive major axis. The second term is due purely to the 
polarisation of the optical trap, for circularly polarised light 
$\phi=\pi/4$ thus maximising the torque transferred to the target particle. 
Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
birefringence, but there are other mechanisms that result in optical torque.

### 2024-10-25 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. This is 
not to say that the scattered field is not influenced by the particle, 
the scattered field can be related to the incident field like so:
\begin{align}
	\label{eq:Rayleigh_scattering}
	\begin{pmatrix}
		{\bf E}^s_\perp \\
		{\bf E}^s_\parallel 
	\end{pmatrix}
	= \begin{pmatrix}
		S_{11} & 0 \\
		0 & S_{22}
	\end{pmatrix}
	\begin{pmatrix}
		{\bf E}^i_\perp \\
		{\bf E}^i_\parallel
	\end{pmatrix}
\end{align}

\noindent
Where $S_{11}$, and $S_{22}$ are elements of the amplitude scattering 
matrix, this is analogous to a single plane wave scattering off a 
sphere. The difference between Eq.~\eqref{eq:Rayleigh_scattering} and
Eq.~\eqref{eq:t_matrix} is that the $T$-matrix accounts for the 
scattering from multiple plane waves that make up the laser. Because the 
particle can be approximated by a point dipole the scattering field is 
linearly related to the scattered field. While not important for optical 
force calculation higher complexity scattering problems can be simplified 
by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}), in which case the scattered field is now 
significant. As the target particle gets larger this assumption fails 
to accurately describe the trapping force \cite{Li2021}. For particle 
sizes close to the laser wavelength the trapping force is best described 
by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle. While one can compute this by directly applying
Lorenz-Mie theory, describing the interactions between focused beams and 
complicated shapes can be computationally slow and memory intensive. As a
result, several alternative scattering approaches have been developed that
allow for faster computation times simplifying the scattering problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}

The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the fields one can derive the 
individual elements of \textbf{T} until a desired accuracy is 
achieved.

For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by 
Mackowski \cite{Mackowski2011} the computational code computes 
the scattered field from each sphere within the target cluster 
and the incident field itself. The scattering between spheres 
will converge to a final result but modern codes truncate 
the calculations to an upper number of iterations (denoted by 
$N_{max}$) to fit a desired accuracy and computational time. 
By centring the expansion at the origin of each sphere the 
individual sphere $T$-matrices can be computed, then using the
addition theorem, the target centred $T$-matrix is computed.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. Each dipole interacts with 
both the incident field and the scattered fields from every other 
dipole surrounding it. The resulting scattered field is identical 
to the scattered field produced by direct integration of Eq.~\eqref{eq:internal} throughout the full particle volume 
\cite{Goedecke1988}. The integral form for the total electric field 
inside a scatterer is given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  \mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

@@

By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps, we can 
discretise the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
noise, as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our noise as a random walk, where at each new time 
step, the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

@@

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

>>>>

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. While 
not important for optical force calculation higher complexity 
scattering problems can be simplified by subdividing the particle 
into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
the scattered field is now significant. As the target particle gets 
larger this assumption fails to accurately describe the trapping force \cite{Li2021}. For larger particles the optical force and scattering 
is better described via the Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce identical results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}

The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the ECBM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix.

For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}) developed by 
Mackowski \cite{Mackowski2011}. Where Lorenz-Mie Theory centres
the spherical harmonics expansion about the sphere's origin it
often struggles to adjust the expansion to any arbitrary origin.
\textit{mstm} utilises an addition theorem to translate the 
individual spherical expansions from the centres of each 
individual sphere to a preset target origin. \textit{mstm}'s 
primary function is to compute the averaged scattering from 
large spherical aggregates, but for our purposes it can be
also utilised to compute $T$ matrices for arbitrary shaped
spherical aggregates.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh Regime, though in order to compute the 
optical force one requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  \mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

@@

By reversing the time step (i.e. use $-\Delta t$) to approximate the
velocity and acceleration based on the previous time steps, we can 
discretise the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
noise, as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our noise as a random walk, where at each new time 
step, the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

@@

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this project we use the work of Vigilante to perform a 
systematic study of the dynamics demonstrated by asymmetric dimers 
in both plane and circularly polarised light. The simulations have 
two primary frames of reference that require explanation. Firstly, 
there is the laboratory frame which describes the coordinate system, 
we define the origin of our simulation as the focus of our trapping 
beam. While the choice in direction for the $x$ and $y$ axis are 
arbitrary the $z$ axis is specifically chosen so that it is aligned 
with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}
